---
title: "MATH 254 - Statistical Modeling and Applications - Lab 2"
author: "Tural Sadigov"
format: 
    pdf:
      theme: 
        light: united
        dark: darkly
editor: visual
chunk_output_type: console
execute:
  out.width: "100%"
  cache: true
---

## NAME: \_\_\_\_\_\_\_\_\_\_\_

## Who are you working with? \_\_\_\_\_\_\_\_\_\_\_

In 2004, the state of North Carolina released to the public a large data set containing information on births recorded in this state. This data set has been of interest to medical researchers who are studying the relation between habits and practices of expectant mothers and the birth of their children. In this mini-project, we will investigate a random sample of 1000 cases from this data using various tools we have been learning from both descriptive and inferential statistics. The data is from Textbook 1 (<https://www.openintro.org/data/index.php?data=ncbirths>) and available on GitHub (<https://github.com/turalsadigov/MATH_254/blob/main/data/ncbirths.csv>).

```{r}
#| message: false
library(tidyverse)

url_remote  <- "https://raw.githubusercontent.com/"
path_github <- "turalsadigov/MATH_254/main/data/"
file_name   <- "ncbirths.csv"
url = paste0(url_remote, path_github, file_name)
nc <- read_csv(url)
nc %>% 
  head(5)
```

#### Descriptive statistics

1.  What are the cases? Categorize variables as numerical variable (discrete or continuous) or categorical variable (regular or ordinal). Mention the unit of each numerical variable. Using command of the data frame in R, also mention the ranges of numerical variables and levels of categorical variables. Example is provided.

    ```{r}
    #| eval: false
    ... %>% 
      glimpse()
    ```

    -   gender: A regular categorical variable with two levels: female and male

2.  Let's look at weights of the babies with different genders and compare them. To get the weights of female babies, one can filter the data with the verb `filter` as below.

    ```{r}
    nc %>% 
      filter(gender == 'female') %>% 
      select(weight) %>% 
      head(5)

    nc %>% 
      filter(gender == 'female') %>% 
      select(weight) %>% 
      summary()
    ```

    Obtain the male weights by modifying the code above in the R chunk below. Create side by side boxplots, and comment on the distribution. Note that, every time you create some graphical summary, you need to have correct labels (with units), titles, etc.

    ```{r}
    #| eval: false

    # extract weights of male babies (add your code below)
    ...

    # get counts for each gender (add your code below)
    ...

    # obtain side by side boxplots (add your code below)
    ...

    ```

3.  Now, let's check if it is plausible that the weights in each category (female, male) are sampled from some hypothetical normal distributions. There are many ways of doing this, but one way is to produce Normal QQ plots which also known as Quantile-Quantile plots. Here, we expect to get a linear pattern if the sample is actually from Normal distribution. If the linearity is not plausible, then one could conclude that normality is not plausible. To get Normal QQ-plot, one can use base R function `qqnorm` or use function `stat_qq` from ggplot. Create side by side QQ plots for the weights of female and male babies. Comment on the scatterplots and poatterns you get. Do you believe that weights of each group are 'sampled' from normal distributions? Why or why not?

    ```{r}
    #| eval: false

    # edit code below
    nc %>% 
      ggplot(aes(sample = ...)) +
      stat_qq() +
      stat_qq_line()+
      facet_wrap(~...)
    ```

4.  Do you think that each sample could be considered as simple random sample? Do you think the samples are independent from each other? Why or why not?

## Inferential Statistics

Assuming the independence within the sample, and independence of samples from each other, and since each data is large enough (sample sizes are greater than 40), we could go ahead with the machinery needed for confidence intervals and hypothesis testing.

### Confidence Intervals

5.  Define parameter of interest (true population means) in the context of the problem. Create 99% confidence interval for difference of true means. Interpret the confidence interval. Is it still possible that, the confidence interval you created does not capture difference of true means (or true difference of population means)? Looking at your confidence interval, is it plausible that there is no difference in true mean weights? Why or why not?

    ```{r}
    #| eval: false

    # load library
    library(infer)

    # calculate and assign observed difference (edit code below)
    observed_difference <- 
      nc %>% 
      specify(weight ~ gender) %>% 
      calculate(stat = '...')

    # print the observed difference
    observed_difference

    # create and assign theoretical sampling distribution of the difference of sample means (edit code below)
    diff_dist <- 
      nc %>% 
      specify(weight ~ gender) %>% 
      assume(...)

    # display the distribution
    diff_dist

    # create and assign confidence interval (edit code below)
    ci <- 
      diff_dist %>% 
      get_confidence_interval(point_estimate = observed_difference,
                              # at the 99% confidence level
                              level = ...,
                              # using the standard error
                              type = "se")

    # print confidence interval
    ci

    # shade the confidence interval (edit code below)
    diff_dist %>%
      visualize() +
      shade_...(endpoints = ci)
    ```

### Hypothesis testing

6.  Let's test if there is a difference between true means (or equivalently, lets check if weights of the babies depends on the gender). State null and alternative hypothesis in the context of the problem and choose significance level.

    $H_0:$

    $H_a:$

    $\alpha = ?$

Since we have checked the conditions for two-sample t test, we go ahead and find test statistic and corresponding p-value (in two different ways -- 'manually' in 7 and 8, and using `infer` in 9 below).

7.  Calculate

    -   Sample mean of female baby weights (let's call it $\bar{x}$)
    -   Sample mean of male baby weights (let's call it $\bar{y}$)
    -   Difference of sample means ($\bar{x}-\bar{y}$)
    -   Sample standard deviation of female baby weights (let's call it $s_x$)
    -   Sample standard deviation of male baby weights (let's call it $s_y$)
    -   Sample size of female baby weights (let's call it $n$)
    -   Sample size of male baby weights (let's call it $m$)
    -   Approximate standard error (deviation) of difference of sample means ($\sqrt{\frac{s_x^2}{n} + \frac{s_y^2}{m}}$)

    ```{r}
    #| eval: false


    # obtain summary statistics: mean, standard deviation and count for each group (edit code below)
    sum_stats <- 
      nc %>% 
      group_by(gender) %>% 
      summarise(ave = ..., 
                st_dev = ...,  
                counts = ...) 

    sum_stats

    x_bar = sum_stats[1,2]
    x_bar
    y_bar = ...
    y_bar
    s_x = ...
    s_x
    s_y = ...
    s_y
    n = ...
    n
    m = ...
    m
    se_difference = sqrt(...)
    se_difference
    ```

8.  Using your results from part (7) above, calculate the test statistic: $$t=\frac{(\bar{x}-\bar{y})-0}{\sqrt{\frac{s_x^2}{n} + \frac{s_y^2}{m}}}$$ Calculate approximate degrees of freedom $df=n+m-2$, and p-value, $2Pr[t_{df}<t]$.

    ```{r}
    #| eval: false

    # edit code below
    df = ...
    df
    t = ...
    t
    p_value = 2*pt(t[1,1], df[1,1])
    p_value 
    ```

9.  Use the following command to find the test statistic, degrees of freedom and p-value again. Check if they matched with above. Interpret the p-value (very carefully).

    ```{r}
    #| eval: false

    # calculate and assign observed t statistic (also test statistic) (edit code below)
    observed_t_stat <- 
      nc %>% 
      specify(...) %>% 
      hypothesise(null = ...) %>% 
      calculate(stat = ...)

    # print the observed t statistic
    observed_t_stat

    # create and assign theoretical sampling distribution of the difference of sample means (edit code below)
    diff_dist <- 
      nc %>% 
      specify(...) %>% 
      assume(...)

    # display the distribution
    diff_dist

    # get p-value (edit code below)
    diff_dist %>% 
      get_p_value(obs_stat = ..., 
                  direction = 'two-sided')

    # shade p-value (edit code below)
    diff_dist %>% 
      visualize() +
      shade_p_value(obs_stat = ..., 
                    direction = ...)
    ```

10. Make a decision in the context of the problem according to the significance level you had chosen above. Would it be still possible that you made an error in the decision? If yes, what kind of error have you made? If an error has been made, then what could be the reason for that?
