---
title: "MATH 254 - Predicting Marriage Happiness"
author:
  - name: Sophie Maniscalco
    email: smanisca@hamilton.edu
    affiliations:
      - name: Hamilton College
  - name: Rachel O'Brien
    email: robrien@hamilton.edu
    affiliations:
      - name: Hamilton College
  - name: Tural Sadigov
    email: tsadigov@hamilton.edu
    affiliations:
      - name: Hamilton College
        orcid: 0000-0002-6446-415X
        address: 198 College Hill Rd
        city: Clinton
        state: NY
        country: USA
        postal-code: 13323-1218
abstract: Due to the high rate of divorce in the United States, a topic of 
  interest for many investigators is relationship happiness in marriages. 
  This project aimed to construct a model predicting relationship happiness using 
  numerous relationship variables. Permutation tests were used to select variables
  that were significantly correlated to relationship happiness -- results indicated 
  that only three out of five potential relationship variables were significantly    
  related to relationship happiness. Seven linear models were constructed using
  the three variables and the best model based on the training-validation slipt 
  and selecting the maximized r-squared value was selected to predict relationship 
  happiness -- specifically, an interaction between hours of sleep per night 
  and argument frequency best predicted relationship happiness. Hours of sleep 
  appeared to serve as a protective factor for couples with higher argument
  frequency, effectively reducing the negative effect of argument frequency
  on relationship happiness.
format:
    html:
      theme: 
        light: united
        dark: darkly
title-block-banner: darkblue
date: October 17, 2022
highlight-style: pygments
chunk_output_type: console
fig-align: center
always_allow_html: true
toc: true
toc-location: right
number-sections: true
page-layout: article
code-overflow: scroll
code-line-numbers: false
code-copy: true
execute:
  echo: true
  warning: false
  eval: true
  output: true
  error: false
  freeze: true
  out.width: "100%"
  cache: true
editor: 
  markdown: 
    wrap: 72
editor_options: 
  chunk_output_type: console
bibliography: references.bib
echo: fenced
code-fold: show
---

# Background and Significance

In 2019, there were 2,015,603 marriages in the United States [@CDC].
This means that marriage is a crucial part of life for a lot of American
people. According to marriage.com, marriage "acts as a social and legal
contract that gives a partner someone to rely on, brings a greater
degree of intimacy and emotional security" @Marriage.com. At least that
this is the ideal goal of marriage. However, also according to the CDC,
in 2019 there were 746,971 divorces in the United States, so this
idealistic goal of marriage is not reached in every relationship.

Given this grim outlook on marriage, we were interested as to what
components of a marriage make it a successful one. Previous studies have
shown that cohabitation before marriage is bad for marriage, with
studies finding on average that couples that cohabited before marriage
had a 33% higher rate of divorce than couples that did not cohabit
before marriage @Atlantic. In addition, marriages where at least one
member is depressed are less positive than marriages where neither
member is depressed @Johnson1997. Additionally, when looking at sleep
and marriage, The American Academy of Sleep Medicine states that "being
stably married or gaining a partner is associated with better sleep in
women than being unmarried or losing a partner" @AASM. With so many
factors that can affect a relationship, how can we predict happiness in
a marriage?

Our main analysis looks at how eleven different variables influence
relationship happiness. We hypothesize that years married, cohabiting
before marriage, hours of sleep a night, days binge drinking, and
argument frequency all impact the happiness of a marriage. Our goal is
to see if it is possible to predict happiness in a marriage through a
combination of these variables.

# Methods

a. *Data collection*. The data, which was collected between 2014 and
2015 only includes couples that "were legally married and had been
living together for a minimum of three years at the time of the study."
The sample was collected using different methods. About 70% of the
same-sex couples in the sample were selected from the Massachusetts
Registry of Vital Records and participants received invitations to
participate in the mail. About 40% of the different-sex couples in the
sample were selected from public records and then asked to participate.
The rest of the participants were recruited through referrals. After all
that, the couples were then asked to refer any other couples that met
the study's requirements. The data itself was collected through a series
of 10-day surveys with questions about their relationship which were
referred to as diaries. The couples were asked to complete these
questionnaires separately. "90% of participants completed all 10 days".
All of this information was provided by icpsr.umich.edu under Health and
Relationships Project, United States, 2014-2015 (ICPSR 37404).

b. *Variable creation*.There are twelve variables we looked at in our
analysis. RELHAPPY, which is our primary variable of interest, is a
discrete numerical variable with range 1 to 7 assessing the respondent's
degree of happiness in the relationship, with higher scores indicating
greater relationship happiness. COHB4MAR is a normal categorical
variable indicating whether or not a couple cohabitated before marriage.
MNCOHB4M is a continuous numerical variable assessing the number of
months that the couple spent cohabiting before marriage if they did
cohabit before marriage, with range 1 to 180. YRSMAR is a continuous
numerical variable assessing how many years a couple has been married
for, with range 0 to 13 years. RAGE is a continuous numerical variable
with range 35 to 65 indicating the respondent's age at the time of the
survey. STRSCALE is a discrete numerical variable with range 9 to 41
indicating how stressful the marriage is; it is an aggregate measure of
multiple variables that were calculated by adding up the scores (from 1
to 7) of the other items. RELREWAR is a discrete numerical variable with
range 1 to 6 representing how rewarding the relationship is, with higher
scores indicating greater feelings that the relationship with their
spouse is rewarding. SPCRITIC is a discrete numerical variable with
range 1 to 5 representing how critical their spouse is, with higher
scores indicating a more critical spouse. ARGUEFRQ is a discrete
numerical variable with range 1 to 5 assessing how often the respondent
argues with their spouse, with higher scores indicating a higher
frequency of arguing. DEPSYM is a discrete numerical variable with range
11 to 38 assessing depressive symptoms in the respondent; this is an
aggregate measure calculated by adding up the positive variable items in
a reverse-coded fashion. CDDAYSBD is a discrete numerical variable with
range 0 to 90 indicating how many days out of the last three months that
the respondent's spouse had four or more drinks. Finally, HRSLEEP is a
continuous numerical variable with range 2 to 10, indicating how many
hours of sleep the respondent gets on an average night. All variables
were taken from Health and Relationships Project, United States,
2014-2015 and, unless otherwise stated, were direct measures.

RELHAPPY1 was created by us in order to better use the data provided.
Originally the data was in string format and we converted it into
numeric. We extracted the numbers provided in the original strings to
create RELHAPPY1.

ARGUEFRQ1 was created by us in order to better use the data provided.
Originally the data was in string format and we converted it into
numeric. We extracted the numbers provided in the original strings to
create ARGUFRQ1.

c. *Analytic Methods*. A bar graph was used to illustrate the
distribution of respondents who cohabited and did not cohabit before
marriage. A histogram was used to illustrate the distribution of daily
hours of sleep received by respondents at night. Five initial
permutation tests were conducted to assess the relationship between the
outcome variable, relationship happiness, and the potential predictor
variables, years married, nightly hours of sleep, number of days in a
three-month period where the partner had more than three drinks, and
argument frequency. Two new variables were constructed for relationship
happiness and argument frequency by extracting the numerical value from
the string data. The relationship between relationship happiness and
cohabitation before marriage was assessed using side-by-side boxplots.

Once we had initial results for our five variables of interest in
relation to relationship happiness, we chose three variables to do
linear regression models with. We chose these variables based on if they
were significantly correlated to relationship happiness or not. Through
our hypothesis testing, we found that two variables, years married and
argument frequency had a significant non-zero correlation coefficient
with relationship happiness. We also chose a third variable, hours of
sleep per night to use in our modeling because it had the lowest p-value
of the variables with a non-significant relationship with relationship
happiness. Then we chose 7 different possible combinations of these
variables to test to see if a linear regression would be a good model
for our data which were, argument frequency, argument frequency added to
years married, argument frequency added to years married, and hours of
sleep per night, argument frequency multiplied by hours of sleep per
night, square root of years married, a degree two polynomial of argument
frequency, and a degree three polynomial of argument frequency. We split
our data into 3 subsections: testing data, training data, and validation
data. Then we used our training data subgroup to create linear
regressions based on our 7 models. Finally, we selected our best model
based on the maximized r-squared value for the validation data. We ended
up selecting our fourth model which was the model that contained
interaction between hours of sleep and argument frequency. Then we used
our testing data to test this model.

# Results

In the results sections we used libraries from the following sources:
@infer, @rlang, @rstudioapi, and @tidyverse.

```{r}
#| message: false
library(tidyverse) 
library(infer)
da37404.0001 <- read_csv('https://raw.githubusercontent.com/turalsadigov/MATH_254/main/Datasets%20for%20projects/happiness_sophie_rachael.csv')
da37404.0001 <- da37404.0001 %>% 
  select(RELHAPPY, YRSMAR, COHB4MAR, HRSLEEP, 
         ARGUEFRQ, CDDAYSBD, MNCOHB4M, STRSCALE, 
         RELWARM, SPCRITIC, DEPSYM) %>% 
  mutate_if(is.character,factor)
# load(file = '37404-0001-Data.rda')
# da37404.0001 <- read_rds(file = '~/Desktop/github_projects/MATH_254/Datasets for projects/happiness_rachael_sophie.RDS')
#load("~/Desktop/github_projects/MATH_254/Projects Submitted/Team_10_Sophie_Rachel/37404-0001-Data.rda")
#da37404.0001 <- tibble::tibble(da37404.0001)
```

## Descriptive statistics

### Descriptive statistics for relationship happiness, our response variable.

Relationship happiness - Descriptive statistics First, we dropped 7 NAs
in the data for a total of 831 observations. From the resulting subset,
the mean relationship happiness was 5.19, with range 1 to 7.

```{r}
data <- da37404.0001 %>% 
  mutate(RELHAPPY1 = str_sub(RELHAPPY, 2, 2)) %>% 
  mutate(RELHAPPY1 = as.numeric(RELHAPPY1))

data %>% 
  drop_na() %>% 
  select(RELHAPPY1) %>% 
  summarise(mean = mean(RELHAPPY1),
            min = min(RELHAPPY1),            
            max = max(RELHAPPY1))
```

### Descriptive statistics for years married.

Years Married - Descriptive statistics No NAs were dropped. The mean
number of years married was 8.69, with range from 0 to 42.5.

```{r}
data %>% 
  drop_na() %>% 
  select(YRSMAR) %>% 
  summarise(mean = mean(YRSMAR), min = min(YRSMAR), max = max(YRSMAR))
```

### Descriptive statistics for cohabitation before marriage.

Cohabitation before marriage - Descriptive statistics No NAs were
dropped. The majority of respondents cohabited before marriage, as
depicted in the bar graph below.

```{r}
data %>% 
  select(COHB4MAR) %>%
  mutate(COHB4MAR1 = str_sub(COHB4MAR, 2, 2)) %>% 
  mutate(COHB4MAR1 = as.numeric(COHB4MAR1)) %>%
  ggplot(aes(COHB4MAR1)) +
  geom_bar() +
  ggtitle('Bar graph of cohabitation before marriage') +
  xlab('Cohabitation before marriage')
```

```{r}
data %>% 
  select(HRSLEEP) %>% 
  summarize(mean = mean(HRSLEEP), 
            minimum = min(HRSLEEP), 
            maximum = max(HRSLEEP))
```

### Descriptive Statistics for hours of sleep.

```{r}
data %>% 
  select(HRSLEEP) %>%
  ggplot(aes(x = HRSLEEP)) +
  geom_histogram(col = 'white') +
  ggtitle('Distribution of Hours of Sleep per Night') +
  xlab('Hours of Sleep per Night')
```

### Descriptive Statistics for argument frequency.

```{r}
data <- data %>% 
  mutate(ARGUEFRQ1 = str_sub(ARGUEFRQ, 2, 2)) %>% 
  mutate(ARGUEFRQ1 = as.numeric(ARGUEFRQ1))

data %>% 
  summarise(mean = mean(ARGUEFRQ1), 
            minimum = min(ARGUEFRQ1), 
            maximum = max(ARGUEFRQ1))
```

### Descriptive Statistics for days in the past three months that spouse consumed more than 4 drinks

```{r}
data %>% 
  select(CDDAYSBD) %>% 
  drop_na() %>% 
  summarize(mean = mean(CDDAYSBD), 
            minimum = min(CDDAYSBD), 
            maximum = max(CDDAYSBD))
  
data %>% 
  select(CDDAYSBD) %>% 
  drop_na() %>% 
  count()
```

## Hypothesis Tests

For the hypothesis tests, we will assess using a significance level of
0.05. Given that we are doing multiple tests and models, we will divide
the significance level by the number of tests and coefficients in our
models to avoid p-hacking. So, for each individual test, the
significance level is 0.0025.

### Permutation test for relation between years married and relationship happiness.

$H_o$: Relationship happiness is not related to years married; the
correlation coefficient is 0.

$H_a:$ Relationship happiness is negatively related to years married:
the correlation coefficient is negative.

Years married vs. relationship happiness from a sub-sample of 831
respondents (after NAs were dropped), the data revealed a significant
negative correlation, based on our p-value, between years married and
relationship happiness, such that as years married increases,
relationship happiness decreases.

```{r}
#| eval: true
correlation_hat <-
  data %>% 
  drop_na() %>% 
  specify(RELHAPPY1 ~ YRSMAR) %>% 
  calculate(stat = 'correlation')

correlation_hat

set.seed(2023)
null_dist <-
  data %>% 
  drop_na() %>% 
  specify(RELHAPPY1 ~ YRSMAR) %>% 
  hypothesize(null = 'independence') %>% 
  generate(reps = 1000, type = 'permute') %>% 
  calculate(stat = 'correlation')

null_dist %>% 
  visualize() +
  shade_p_value(obs_stat = correlation_hat,
                direction = "two-sided")

null_dist %>% 
  get_p_value(obs_stat = correlation_hat,
              direction = 'two-sided')
```

### Preliminary result of relationship between months cohabiting before marriage and relationship happiness.

These boxplots provide preliminary results that the tail ends of
happiness (i.e., perfect relationship happiness and extreme unhappiness)
correspond to longer cohabitation periods before marriage, as indicated
by the IQR of the boxplots.

```{r}
da37404.0001 %>% 
  select(MNCOHB4M, RELHAPPY) %>% 
  drop_na() %>% 
  ggplot(aes(x = MNCOHB4M, y = RELHAPPY)) +
  geom_boxplot() +
  ggtitle('Happiness with Relationship and Pre-Marriage Cohabitation') +
  xlab('Months Cohabitated before Marriage') +
  ylab("Happiness with relationship")
```

### Permutation test for relation between cohabitation before marriage and relationship happiness.

$H_o$: Relationship happiness is not related to cohabitation before
marriage; the correlation coefficient is 0.

$H_a:$ Relationship happiness is negatively related to cohabitation
before marriage; the correlation coefficient is negative.

Cohabitation before marriage vs. relationship happiness from a
sub-sample of 831 respondents (after NAs were dropped). The boxplot for
respondents that cohabited before marriage has a smaller inter-quartile
range than the boxplot for respondents who did not cohabit before
marriage, indicating that couples that cohabited before marriage are
generally happier with their relationship. This finding is contrary to
previous findings that suggest cohabiting before marriage is bad for the
relationship. However, follow-up hypothesis testing using a permutation
test for difference in means revealed that there was a non-significant
difference in relationship happiness between couples that cohabited and
those that did not cohabit before marriage.

```{r}
data %>% 
  select(RELHAPPY1, COHB4MAR) %>% 
  drop_na() %>% 
  ggplot(aes(COHB4MAR, RELHAPPY1)) +
  geom_boxplot() +
  xlab('Cohabitation before marriage') +
  ylab('Relationship happiness')
```

```{r}
obs_stat <-
  data %>% 
  specify(RELHAPPY1 ~ COHB4MAR) %>% 
  calculate(stat = 'diff in means')
obs_stat

set.seed(2023)
null_dist <-
  data %>% 
  specify(RELHAPPY1 ~ COHB4MAR) %>% 
  hypothesize(null = 'independence') %>% 
  generate(reps = 1000, type = 'permute') %>% 
  calculate(stat = 'diff in means')

null_dist %>% 
  visualize() +
  shade_p_value(obs_stat = correlation_hat,
                direction = "two-sided")

null_dist %>% 
  get_p_value(obs_stat = correlation_hat,
              direction = 'two-sided')
```

### Permutation Test for hours of sleep per night and relationship happiness.

Hypothesis Testing for independence between hours of sleep per night and
relationship happiness

Null Hypothesis: hours of sleep per night and relationship happiness are
independent. The correlation coefficient is 0.

Alternative Hypothesis: hours of sleep per night and relationship
happiness are not independent. The correlation coefficient is not 0.

```{r}
#| message: false
#| warning: false
obs_stat <- 
  data %>% 
  specify(RELHAPPY1 ~ HRSLEEP) %>% 
  calculate(stat = 'correlation')

set.seed(2022)
null_dist <- 
  data %>% 
  specify(RELHAPPY1 ~ HRSLEEP) %>% 
  hypothesize(null = 'independence') %>% 
  generate(reps = 1000, type = 'permute') %>% 
  calculate(stat = 'correlation')

null_dist %>% 
  visualize()

p_value <- 
  null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = 'both')
p_value

null_dist %>% 
  ggplot(aes(stat)) +
  geom_density(fill = 'blue', alpha = 0.5)  +
  geom_vline(xintercept = obs_stat$stat, 
             color = "red", 
             size=1.5) +
  geom_vline(xintercept = - obs_stat$stat, 
             color = "red", 
             size=1.5)
obs_stat
```

This permutation test results in a p-value of 0.004, which is slightly
larger than our significance level of 0.0025. Therefore we fail to
reject the null-hypothesis that the correlation coefficient between
relationship happiness and hours of sleep per night is 0.

### Permutation test for independence between days partner consumed more than 4 drinks in the last 90 days and relationship happiness:

Null Hypothesis: Days partner consumed more than 4 drinks in the last 40
days and relationship happiness are independent. The correlation
coefficient is 0.

Alternative Hypothesis: Days partner consumed more than 4 drinks in the
last 40 days are not independent. The correlation coefficient is not 0.

```{r}
obs_stat <- 
  data %>% 
  specify(RELHAPPY1 ~ CDDAYSBD) %>% 
  calculate(stat = 'correlation')

set.seed(2022)
null_dist <- 
  data %>% 
  specify(RELHAPPY1 ~ CDDAYSBD) %>% 
  hypothesize(null = 'independence') %>% 
  generate(reps = 1000, type = 'permute') %>% 
  calculate(stat = 'correlation')

null_dist %>% 
  visualize()

p_value <- 
  null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = 'both')
p_value

null_dist %>% 
  visualize() +
  theme_dark() +
  shade_p_value(obs_stat = obs_stat, direction = 'both')

null_dist %>% 
  ggplot(aes(stat)) +
  geom_density(fill = 'blue', alpha = 0.5)  +
  geom_vline(xintercept = obs_stat$stat, 
             color = "red", 
             size=1.5) +
  geom_vline(xintercept = - obs_stat$stat, 
             color = "red", 
             size=1.5)
obs_stat
```

This permutation test results in a p-value of 0.566 which is greater
than our significance level of 0.0025, so we fail to reject the null
hypothesis that days partner consumed more than 4 drinks in the last 40
days is independent from relationship happiness.

### Permutation test for independence between argument frequency and relationship happiness:

Null Hypothesis: Argument frequency and relationship happiness are
independent. The correlation coefficient is 0.

Alternative Hypothesis: argument frequency and relationship happiness
are not independent. The correlation coefficient is not 0.

```{r}
#| warning: false

obs_stat <- 
  data %>% 
  specify(RELHAPPY1 ~ ARGUEFRQ1) %>% 
  calculate(stat = 'correlation')

set.seed(2022)
null_dist <- 
  data %>% 
  specify(RELHAPPY1 ~ ARGUEFRQ1) %>% 
  hypothesize(null = 'independence') %>% 
  generate(reps = 1000, type = 'permute') %>% 
  calculate(stat = 'correlation')

null_dist %>% 
  visualize()

p_value <- 
  null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = 'both')
p_value

null_dist %>% 
  visualize() +
  theme_dark() +
  shade_p_value(obs_stat = obs_stat, direction = 'both')

null_dist %>% 
  ggplot(aes(stat)) +
  geom_density(fill = 'blue', alpha = 0.5)  +
  geom_vline(xintercept = obs_stat$stat, 
             color = "red", 
             size=1.5) +
  geom_vline(xintercept = - obs_stat$stat, 
             color = "red", 
             size=1.5)
obs_stat
```

This permutation test results in a p-value of 0 which is less than our
significance level of 0.0025 which leads us to reject the null
hypothesis that the correlation coefficient between argument frequency
and relationship happiness is 0. So there is a significant correlation
between the argument frequency and relationship happiness.

### New variable ARGUEFRQ1

The variable ARGUEFRQ1 was constructed by taking the second character in
the ARGUEFRQ strings in the data and mutating them into a new column
that is a numeric variable so that we could more easily use the data for
testing.

```{r}
data %>% 
  select(ARGUEFRQ1) %>% 
  summarize(mean = mean(ARGUEFRQ1))
```

## Linear Model

For our linear model, we used libraries from the following sources:
@tidymodels, @tidyverse, @ggfortify, and @vip.

```{r}
library(tidymodels)
library(tidyverse)
library(ggfortify)
library(vip)

df <- data %>% 
  select(RELHAPPY1, ARGUEFRQ1, HRSLEEP, YRSMAR) %>% 
  drop_na()

set.seed(1800)
df_split <- initial_split(data = df, prop = 0.80, strata = RELHAPPY1)
df_split

df_not_testing <- training(df_split)
df_testing <- testing(df_split)


set.seed(123)
df_split_2 <- initial_split(data = df_not_testing,
                            prop = 0.75,
                            strata = RELHAPPY1)

df_training <- training(df_split_2)
df_validation <-testing(df_split_2)

model_specs <-
  linear_reg() %>% 
  set_engine('lm') %>% 
  set_mode('regression')

model_1 <-
  model_specs %>% 
  fit(RELHAPPY1 ~ ARGUEFRQ1, data = df_training)

model_2 <-
  model_specs %>% 
  fit(RELHAPPY1 ~ ARGUEFRQ1 + YRSMAR, data = df_training)

model_3 <-
  model_specs %>% 
  fit(RELHAPPY1 ~ ARGUEFRQ1 + YRSMAR + HRSLEEP, data = df_training)

model_4 <-
  model_specs %>% 
  fit(RELHAPPY1 ~ ARGUEFRQ1*HRSLEEP, data = df_training)

model_5 <-
  model_specs %>% 
  fit(RELHAPPY1 ~ sqrt(YRSMAR), data = df_training)

model_6 <-
  model_specs %>% 
  fit(RELHAPPY1 ~ poly(ARGUEFRQ1, degree = 2, raw = TRUE), data = df_training)

model_7 <-
  model_specs %>% 
  fit(RELHAPPY1 ~ poly(ARGUEFRQ1, degree = 2, raw = TRUE) + YRSMAR, data = df_training)

fitted_models <- list(model_1, model_2, model_3,
                      model_4, model_5, model_6,
                      model_7)

# function for validation performance
validation_performance <- function(model){
  model_results <-
    df_validation %>% 
    bind_cols(predict(model, df_validation)) %>% 
    select(RELHAPPY1, .pred)
  my_metrics <- metric_set(rmse, rsq, mae)
  output <-
    model_results %>% 
    my_metrics(RELHAPPY1, .pred) %>% 
    select(-.estimator) %>% 
    pivot_wider(names_from = .metric, values_from = .estimate)
  return(output)
}

validation_results <- tibble()
for(model in fitted_models){
  validation_results <-
    validation_results %>% 
    bind_rows(validation_performance(model))
}
validation_results

# CHOOSE A MODEL and do a last fit!

model_4_updated <-
  model_specs %>% 
  fit(RELHAPPY1 ~ ARGUEFRQ1*HRSLEEP, data = df_not_testing)

tidy(model_4_updated)

# report the following test result

test_results <-
  df_testing %>% 
  bind_cols(predict(model_4_updated, new_data = df_testing)) %>% 
  select(RELHAPPY1, .pred)

my_metrics <- metric_set(rmse, rsq, mae)

test_results %>% 
  my_metrics(RELHAPPY1, .pred)
```

# Discussion/Conclusions

Our objective in this project was to find a way to predict happiness
within a marriage using a combination of factors including, hours of
sleep per night, years married, cohabitation before marriage, argument
frequency, and drinking habits. The best model that we found is $$
RELHAPPY1 = 
$$ $$7.29 - 1.01ARGUEFRQ1 - 0.1404HRSLEEP +
0.0787ARGUEFRQ1*HRSLEEP.
$$ The $R^2$ value for this model is .193. This means that around 19% of
the variance in marriage relationship happiness can be explained by our
model. Our model suggests that a higher frequency of arguments predicts
a lower rate of happiness in a relationship. However, if the couple
sleeps more hours per night, the argument frequency has less of a
negative effect on happiness. Previous studies have shown that
cohabitation before marriage is bad for marriage @Atlantic, however, we
did not find a significant relationship between the relationship
happiness and cohabitation before marriage. Additionally, we found that
hours of sleep has a positive effect on a marriage which aligns with The
American Academy of Sleep Medicine's study on sleep and relationships
@AASM. Happiness is hard to find and define in real life, and it is also
hard to predict, so there are some limitations we have to consider. Our
data set contained around 500 variables, and we did not have the
capacity to use all of them. Additionally, many of the variables we did
not choose had a high population of N/A inputs which is one reason they
did not end up on our final list of variables. Since we only selected
five possible explanatory variables, there were a lot of other factors
that could lead to relationship happiness including, age, political
beliefs, religion, and other variables. If possible future research
could look at more models with different variables by using imputation
techniques to use variables with a lot of missing values.

# Appendix

In the appedix we used the libraries @tidyverse and @infer.

## Loading Libraries and Data File.

```{r}
#| eval: false
library(tidyverse)
library(infer)
load(file = '37404-0001-Data.rda')
```

## Visualizing cohabitation before marriage.

Visualizing breakdown of cohabitation before marriage - COHAB4MAR This
variable is a non-ordinal categorical variable with two levels that
tells whether or not the couple cohabited before marriage. Seems like
the vast majority of couples cohabit before marriage.

```{r}
da37404.0001 %>% 
  ggplot(aes(x = COHB4MAR)) +
  geom_bar() +
  ggtitle('Bar plot depicting cohabitation before marriage') +
  xlab('Cohabitation before marriage')

```

## Visualizing months cohabiting before marriage.

Visualizing number of months cohabiting before marriage - MNCOHB4M This
is a continuous numerical variable with range 0 to infinity that tells
us the number of months that the couple spent cohabiting before
marriage. Problem: if couple did not cohabit before marriage, data was
entered as NA rather than 0. Will have to mutate the variable.

```{r}
da37404.0001 %>% 
  ggplot(aes(x = MNCOHB4M)) +
  geom_histogram() +
  ggtitle('Histogram of months cohabited prior to marriage') +
  xlab('Number of Months')
```

## Visualizing years married.

Visualizing years married among couples in this data set - YRSMAR. This
is a continuous numerical variable with range 0 to infinity that tells
us how long a couple has been married at the time of their interview.

```{r}
da37404.0001 %>% 
  select(YRSMAR) %>% 
  ggplot(aes(x = YRSMAR)) +
  geom_histogram() +
  xlab('Years married')
```

## Visualizing stress levels.

Visualizing stress scores - STRSCALE This is a continuous numerical
variable with range 0 to 41, indicating the sum of stress related
question scores. Higher numbers correspond to higher stress levels.

```{r}
da37404.0001 %>% 
  select(STRSCALE) %>% 
  ggplot(aes(STRSCALE))+
  geom_histogram() +
  xlab('Stress scale')

da37404.0001 %>% 
  summarise(range = range(STRSCALE))
```

## Visualizing relationship happiness, our response variable.

Visualizing relationship happiness - RELHAPPY This is a ordinal
categorical variable with range 1 to 7.

```{r}
da37404.0001 %>% 
  select(RELHAPPY) %>% 
  ggplot(aes(RELHAPPY)) +
  geom_density() +
  theme(axis.text.x = element_text(angle = 90)) + 
  xlab('Relationsip Happiness')
```

## Visualizing relationship warmness.

Visualing relationship warmness - RELRWAR This is a ordinal categorical
variable with range 1 to 7.

```{r}
da37404.0001 %>% 
  select(RELWARM) %>% 
  ggplot(aes(RELWARM)) +
  geom_density() +
  theme(axis.text.x = element_text(angle = 90)) +
  xlab('Relationship Warmness')
```

## Visualizing how critical one's spouse is.

Visualizing how critical spouse's partner is - SPCRICTIC This is a
ordinal categorical variable with range 1 to 5.

```{r}
da37404.0001 %>% 
  select(SPCRITIC) %>% 
  ggplot(aes(SPCRITIC)) +
  geom_density() +
  xlab('Criticalness of Spouse')
```

## Visualizing argument frequency.

ARGUEFRQ: An ordinal categorical variable with how often the couple
argues with 5 levels, (1) Very Rarely, (2) Rarely, (3) Sometimes, (4)
Often, (5) Very Often. This variable gives us information about how
frequently the couple argues.

```{r}
da37404.0001 %>% 
  select(ARGUEFRQ) %>% 
  ggplot(aes(x = ARGUEFRQ)) +
  geom_bar() +
  ggtitle("Frequency of arguements in couples") +
  xlab("Frequency of arguements")
```

## Visualizing depression symptoms.

DEPSYM: A discrete numerical variable range 11 to 38 that tells us about
how many depressive symptoms the respondant faces.

Note: This variable was created by adding up the ranking (1-5) of other
categories to determine the overall depressive symptoms numerical value.

```{r}
da37404.0001 %>% 
  select(DEPSYM) %>% 
  summarise(range = range(DEPSYM))
```

```{r}
da37404.0001 %>% 
  select(DEPSYM) %>% 
  ggplot(aes(x = DEPSYM)) +
  geom_boxplot() +
  xlab('Depression symptoms')
```

## Visualizing days in which spouse has had more than four drinks in one day in the last three months

CDDAYSBD: A continuous numerical variable with range 0 to 90 which tells
us how many days has the spouse had 4+ drinks in the last 3 months.

Problem: There are some N/As.

```{r}
da37404.0001 %>% 
  select(CDDAYSBD) %>% 
  drop_na() %>% 
  summarise(range = range(CDDAYSBD))
```

```{r}
da37404.0001 %>% 
  select(CDDAYSBD) %>% 
  drop_na() %>% 
  ggplot(aes(x = CDDAYSBD)) +
  geom_histogram() +
  xlab('Days partner had more than 3 drinks in last three months')
```

## Visualizing hours of sleep.

HRSLEEP: A continuous numerical variable with range 2 to 10 that tells
us how many hours of sleep per/day the respondent gets.

```{r}
da37404.0001 %>% 
  select(HRSLEEP) %>% 
  summarise(range = range(HRSLEEP))
```

```{r}
da37404.0001 %>% 
  select(HRSLEEP) %>% 
  drop_na() %>% 
  ggplot(aes(x = HRSLEEP)) +
  geom_density(fill = 'blue', alpha = 0.5) +
  xlab('Hours of sleep')
```

# References
