---
title: "NBA Salaries for 2019-2020"
author: "Tural Sadigov"
date: today
format: 
    html:
      theme:
        light: united
        dark: superhero
editor: visual
---

## Load libraries

```{r}
library(tidyverse)
```

Load data.

```{r}
nba = read_table(file = 'nba_salaries_text')
```

Extract the column as a vector/array.

```{r}
#| out.width: "100%"

# pull it
nba_salaries <- 
  nba %>% 
  pull(Salary)
# display first 30 of them
nba_salaries %>% 
  head(30)
```

Distribution.

```{r}
#| out.width: "100%"
nba %>% 
  ggplot(aes(x = Salary)) +
  geom_histogram(aes(y = ..density..),
                 color = 'white', 
                 fill = 'blue', 
                 alpha = 0.5) +
  geom_density(lwd = 2) +
  xlab('NBA salaries (in million $)') +
  ggtitle('Distribution of NBA salaries - 2019/20 season')
```

Numerical summaries.

```{r}
#| out.width: "100%"

# summaries
nba %>% 
  summarise(mean = mean(Salary), 
            median = median(Salary), 
            sd = sd(Salary), 
            var = var(Salary),
            iqr = IQR(Salary))
```

We will treat this salary data as our population since it has all the salaries for all NBA players from 2019-2020 season. Now, what if we choose 10 players out of `r dim(nba)[1]` players, and find the sample mean. How good would that be an estimate of the true mean, `r mean(nba$Salary)`? Lets do an experiment. We choose a sample without replacement, and calculate the sample mean.Do it many many time.

```{r}
set.seed(2022)
nba %>% 
  sample_n(size = 10, 
           replace = FALSE) %>% 
  summarise(mean = mean(Salary))
```

Lets do this many many times and keep records of each sample mean. One can do this with a **for** loop, but we will use a package named **`infer`** to abstract away the process.

```{r}
library(infer)
nba %>% 
  rep_sample_n(size = 10, 
               reps = 2, # play with number of repetitions
               replace = FALSE) 
```

Lets group them by replication, and then find average salary for each replication.

```{r}
set.seed(2022)
nba %>% 
  rep_sample_n(size = 10, 
               reps = 500,
               replace = FALSE) %>% 
  group_by(replicate) %>% 
  summarise(mean = mean(Salary))
```

Now, we can get sampling distribution for 500 repeated samples.

```{r}
#| out.width: "100%"

set.seed(2022)
nba %>% 
  rep_sample_n(size = 10, 
               reps = 500,
               replace = FALSE) %>% 
  group_by(replicate) %>% 
  summarise(mean = mean(Salary)) %>% 
  ggplot(aes(x = mean)) +
  geom_histogram(aes(y = ..density..), 
                 color = 'white',
                 fill = 'blue', 
                 alpha = 0.5) +
  geom_density(lwd = 2)
```

Try the sampling distribution for 10,000 samples!

```{r}
#| out.width: "100%"

set.seed(2022)
nba %>% 
  rep_sample_n(size = 10, 
               reps = 10000,
               replace = FALSE) %>% 
  group_by(replicate) %>% 
  summarise(means = mean(Salary)) %>% 
  ggplot(aes(x = means)) +
  geom_histogram(aes(y = ..density..), 
                 color = 'white',
                 fill = 'blue', 
                 alpha = 0.5) +
  geom_density(lwd = 2)
```

Lets approximate mean and standard distribution of the sampling distribution.

```{r}
#| out.width: "100%"

set.seed(2022)
resampling_data <- 
  nba %>% 
  rep_sample_n(size = 10, 
               reps = 10000,
               replace = FALSE) %>% 
  group_by(replicate) %>% 
  summarise(means = mean(Salary))
resampling_data
```

Compare mean of the population and mean of means!

```{r}
mean(nba$Salary)
mean(resampling_data$means)
```

Compare standard deviation of the population and standard deviation of means!

```{r}
sd(nba$Salary)
sd(resampling_data$means)
```

It turns out that one could calculate the mean and the standard deviation of the sampling distribution via the following formulas:

$$
\mu_{\bar{X}} = \mu_X, \quad\quad \sigma_{\bar{X}} = \frac{\sigma_X}{\sqrt{n}}
$$

Lets check this.

```{r}
sd(nba$Salary)/sqrt(10)
sd(resampling_data$means)
```

What if we would like to create/investigate/understand the sampling distribution of the sample maximum?

```{r}
#| out.width: "100%"

set.seed(2023)
nba %>% 
  rep_sample_n(size = 10, 
               reps = 10000,
               replace = FALSE) %>% 
  group_by(replicate) %>% 
  summarise(max = max(Salary)) %>% 
  ggplot(aes(x = max)) +
  geom_histogram(aes(y = ..density..), 
                 color = 'white',
                 fill = 'blue', 
                 alpha = 0.5) +
  geom_density(lwd = 2)
```
