---
title: "Chernobyl data - Statistical Inference using infer package"
author: "Tural Sadigov"
date: today
format: 
    html:
      theme: 
        light: united
        dark: darkly
editor: visual
chunk_output_type: console
execute:
  out.width: "100%"
  cache: true
---

## Libraries

```{r}
#| message: false
library(tidyverse)
library(infer)
```

Example from class says that out of 244 birds, 16% had white feather. Lets simulate such data.

```{r}
n = 0.16*244
feather <- rep('white_feather', n)
no_feather <- rep('no_white_feather', 244-n)
set.seed(2022)
chernobyl <- sample(c(feather, no_feather))
chernobyl[1:10]

chernobyl_df <- tibble(chernobyl)
chernobyl_df
```

## Sanity check - 1

```{r}
chernobyl_df %>% 
  count(chernobyl)
```

## Sanity check - 2

```{r}
p_hat <-   
  chernobyl_df %>% 
  specify(response = chernobyl, 
          success = 'white_feather') %>% 
  calculate(stat = 'prop')
p_hat
```

## Observed test statistics (z)

```{r}
z_stat<- 
  chernobyl_df %>% 
  specify(response = chernobyl, 
          success = 'white_feather') %>% 
  hypothesize(null = 'point', 
              p = .02) %>% 
  calculate(stat = 'z')
z_stat
```

## Sanity check - 3

$$
z = \frac{\hat{p}- p}{\sqrt{\frac{p(1-p)}{n}}}
$$

```{r}
(p_hat - 0.02)/(sqrt(0.02*0.98/244))
```

## Check the assumptions of proportion test

1.  Simple random sample: birds are caught independently from each other from Chernobyl area.

2.  $\frac{n}{N} = \frac{244}{1000000} = 0.000244<<0.1$

3.  $n = 244 > 40$

## Application of hypothesis testing

-   Null Hypothesis vs Alternative Hypothesis, Significance level

    $H_0:$ True proportion of the birds with white feather in Chernobyl region is 0.02, i.e., $p = 0.02$.

    $H_a:$ True proportion of the birds with white feather in Chernobyl region is larger than 0.02, i.e., $p > 0.02$.

    $\alpha = 0.001$

-   Sampling Distribution under the null hypothesis with checked assumptions.

    $\hat{p}_n$ is approximately normal with $\mu = p$ and $\sigma = \sqrt{\frac{p(1-p)}{n}}$.

```{r}
# approximated sampling distribution of p-hat under the checked assumptions
z_dist <- 
  chernobyl_df %>% 
  specify(response = chernobyl, 
          success = 'white_feather') %>% 
  hypothesize(null = 'point', 
              p = 0.02) %>% 
  assume(distribution = 'z') 
z_dist %>% 
  visualize()
```

-   Observed statistic, test statistic , p-value

### Observed statistic, test statistic, p-value

```{r}
p_hat
z_stat
z_dist %>% 
  get_p_value(obs_stat = z_stat, 
              direction = 'right')
```

### Visualization

```{r}
z_dist %>% 
  visualize() +
  shade_p_value(z_stat, direction = "right")
```

-   Decision:

With the p-value that is nearly 0 (thus, lower than $\alpha = 0.005$),

-   we reject the Null hypothesis that true proportion of birds with white feather is 0,02

-   observed statistic of 16% cannot be attributed to chance alone.

We are NOT claiming that nuclear reactor disaster is the CAUSE of the observed statistic. We are just saying that observed effect cannot be reasonably attributed to chance alone.


