---
title: "County data - Statistical Inference using infer package"
author: "Tural Sadigov"
date: today
format: 
    html:
      theme: 
        light: united
        dark: darkly
editor: visual
chunk_output_type: console
execute:
  out.width: "100%"
  cache: true
---

## Load libraries and R data

```{r}
#| message: false

# our new universe
library(tidyverse)

# reading from CSV
remote <- 'https://raw.githubusercontent.com/'
my_github <- 'turalsadigov/'
path <- 'MATH_254/main/data/county.csv'
url <- paste0(remote, my_github, path)
county = read_csv(url)

county %>% 
  head(3)
```

## Change the data format to tibble (advanced data frame) and view portion of it.

```{r}
county <- as_tibble(county)
county
```

## Sample from the population!

```{r}
set.seed(123)
county_small <- 
  county %>% 
  slice_sample(n = 100)
county_small
```

## Boxplots for poverty with each level of metro (NAs dropped).

```{r}
#| out.width = "100%"
# boxplot - facet wrapped
county_small %>% 
  drop_na() %>% # drop nas
  ggplot(aes(x = poverty)) +
  geom_boxplot(fill = 'green', alpha = 0.5) +
  facet_wrap(~metro) + # facet wrap with metro levels
  coord_flip() # make them vertical
```

## Can we facet-wrap it?

```{r}
#| out.width = "100%"
# distribution - histogram + facet wrap
county_small %>% 
  drop_na() %>% 
  ggplot(aes(x = poverty)) +
  geom_density(fill="lightblue") +
  facet_wrap(~metro)
```

## We can do better.

```{r}
#| out.width = "100%"
# distribution - histogram + facet wrap
county_small %>% 
  drop_na() %>% 
  ggplot(aes(x = poverty, fill= metro)) +
  geom_density(alpha = 0.4) 
```

```{r}
county_small %>% 
  drop_na() %>% 
  group_by(metro) %>% 
  summarise(means = mean(poverty))
```

## Hypothesis testing with `infer`

```{r}
county_small <- 
  county_small %>% 
  drop_na()
```

```{r}
library(infer)
observed_diff <- 
  county_small %>% 
  specify(response = poverty, explanatory = metro) %>% 
  calculate(stat = 'diff in means')
observed_diff
```

So can the observed mean difference, `r observed_diff`, be attributed to chance alone, or is there a relationship between these two variables that could account for the difference in mean poverty percentages?

## Sanity check

```{r}
county_small %>% 
  specify(response = poverty, explanatory = metro) %>% 
  hypothesize(null = 'independence') %>% 
  calculate(stat = 'diff in means')
```

## Check the assumptions of difference of means test

1.  Simple random samples: counties are randomly selected from all 2560 US counties.

2.  $\frac{n}{N} = \frac{100}{2560} = 0.0391<<0.1$

3.  $n = 100>40$

## Application of hypothesis testing

-   Null Hypothesis vs Alternative Hypothesis, Significance level

    $H_0:$ The difference of true means poverty percentage between counties that has no metropolitan city in them and the counties that has the metropolitan city in them is 0 (i.e., poverty level of a county is independent of whether there is a metropolitan city in it or not).

    $H_a:$ The difference of true means poverty percentage between counties that has no metropolitan city in them and the counties that has the metropolitan city in them positive (i.e., poverty level of a county DOES depend on whether there is a metropolitan city in it or not).

    $\alpha = 0.01$

-   Sampling Distribution under the null hypothesis with checked assumptions. Under the null hypothesis of no difference,

    $\frac{\bar{x}-\bar{y}}{\sigma_{\bar{x}-\bar{y}}}$ has approximately t-distribution with some degrees of freedom.

    ```{r}
    t_dist <- 
      county_small %>% 
      specify(response = poverty, 
              explanatory = metro) %>% 
      hypothesize(null = 'independence') %>% 
      assume(distribution = 't') 
    t_dist %>% 
      visualize()
    ```

-   Observed statistic, test statistic, p-value

    ```{r}
    observed_diff
    t_stat <- 
      county_small %>% 
      specify(response = poverty, 
              explanatory = metro) %>% 
      hypothesize(null = 'independence') %>% 
      calculate(stat = 't')
    t_stat
    p_value <- 
      t_dist %>% 
      get_p_value(obs_stat = t_stat, 
                  direction = 'right')
    p_value
    ```

### Visualization

```{r}
t_dist %>% 
  visualize() +
  shade_p_value(t_stat, direction = "right")
```

## Decision

With the p-value of `r p_value` that is not small enough (less than our pre-determined significance level of 0.01),

-   we FAIL to reject the Null hypothesis that the difference between true averages poverty percentages between counties that have no metropolitan city in them and the counties that have the metropolitan city in them is 0 (i.e., we FAIL to reject that poverty level of a county is independent of whether there is a metropolitan city in it or not).

-   observed statistic of the difference of 3.15% can be attributed to chance alone, and there is no strong evidence against it.
