---
title: "Queries in dplyr and Permutation Test for slopes in the Simple Linear Regression"
author: "Tural Sadigov"
date: today
format: 
    html:
      theme: 
        light: united
        dark: darkly
editor: visual
chunk_output_type: console
fig-align: center
always_allow_html: true
toc: true
toc-location: right
number-sections: true
page-layout: article
code-overflow: scroll
code-line-numbers: false
code-copy: true
execute:
  echo: true
  warning: false
  eval: true
  output: true
  error: false
  freeze: true
  out.width: "100%"
  cache: true
---

## libraries

```{r}
library(tidyverse)
library(infer)
library(stats2data)
```

## dplyr verbs

-   select

-   mutate

-   group_by

-   summarise

-   filter

-   arrange

## queries

### query 1

Extract/list/obtain state name, homeownership and unemployment rates for counties from the state of indiana or Illionois with unemployment rate below 3%. Arrange them by descending order of homeownership. 

```{r}
stats2data::county %>% 
  select(state, homeownership, unemployment_rate) %>% 
  filter(state == "Indiana" | state == 'Illinois') %>% 
  filter(unemployment_rate < 3) %>% 
  arrange(desc(homeownership))
```

### query 2

Extract/list/obtain state name,  unemployment rates and metro variable for counties from the state of indiana or Illionois with unemployment rate below 3%. arrange them by metro variable where homeownership is descending within levels of metro.

```{r}
stats2data::county %>% 
  filter(unemployment_rate < 3) %>% 
  select(state, homeownership, metro) %>% 
  filter(state == "Indiana" | state == 'Illinois') %>% 
  arrange(metro, desc(homeownership))
```

### query 3

Extract all NY counties where there is a NO metropolitan city within the county.Print all rows with four columns: state, metro, poverty, homeownership.

```{r}
stats2data::county %>% 
  filter(state == 'New York' & metro == 'no') %>% 
  select(state, metro, poverty, homeownership) %>% 
  print(n = 24)
```

## inference on a subgroup

Extract all NY counties where there is a NO metropolitan city within the county, and test if slope of the simple linear regression between homeownership versus poverty is not 0 for these counties (i.e., check if there is a dependency).

### parameter of interest

True slope below: $\beta_1$.


$$
H = \beta_0 + \beta_1 p + Z
$$


### hypothesis and significance level.

$H_0$: True slope in the simple linear regression model above, $\beta_1$, is 0 (i.e., homeownership does not depend on poverty for NY counties that have no metropolitan city within them).

$H_a$: True slope in the simple linear regression model above, $\beta_1$, is NOT 0 (i.e., homeownership depends on poverty for NY counties that have no metropolitan city within them).

$\alpha = 0.001$

### permutation Test

Null hypothesis suggest independence between two numerical variables. We shuffle/permute one and  keep the other fixed to obtain a permutation distribution for $\hat{beta_1}$, and this permutation distribution will approximate true sampling distribution of $\hat{beta_1}$ under Null hypothesis of independence.

### observed slope, null distribution and p-value

```{r} 
df <- 
  stats2data::county %>% 
  filter(state == 'New York' & metro == 'no') %>% 
  select(state, metro, poverty, homeownership)

obs_stat <- 
  df %>% 
  specify(poverty ~ homeownership) %>% 
  calculate(stat = 'slope')

obs_stat


set.seed(123)
null_dist <- 
  df %>% 
  specify(poverty ~ homeownership) %>% 
  hypothesise(null = 'independence') %>% 
  generate(reps = 1000, type = 'permute') %>% 
  calculate(stat = 'slope')

null_dist %>% 
  visualise()

null_dist %>% 
  get_p_value(obs_stat = obs_stat, direction = 'both')

null_dist %>% 
  visualise() +
  shade_p_value(obs_stat = obs_stat, direction = 'both')
```

Scatterplot

```{r}
df %>% 
  ggplot(aes(x = poverty, y = homeownership)) +
  geom_point(size = 3) +
  geom_jitter() +
  geom_smooth(method = "lm", lwd = 2)
```


### Decision

Since p-value of $0.034$ > $\alpha = 0.001$, we conclude that there is no strong evdience against the assumption of indepdence between poverty and homeownership in rural counties of NY, i.e., it could very well be the case that $\beta_1 = 0$, and observed slope of `r obs_stat` could be due to random chance and is not statistically significant. 