---
title: "MATH 254 - Statistical Modeling and Applications - HW0 "
output: pdf
execute: 
  eval: false
  echo: true
---

## Download R and R Studio

This homework is to ensure that you have R Studio running on your personal computer. If you do not have a personal computer, I believe, LITS can help you to get one.

1.  Download and install R.

    -   For Windows: [https://cran.r-project.org/bin/windows/base/](https://cran.r-project.org/bin/windows/base/){target="_blank"}

    -   For Mac: [https://cran.r-project.org/bin/macosx/](https://cran.r-project.org/bin/macosx/){target="_blank"}

```{=html}
<!-- -->
```
2.  Download and install R Studio:
    -   [https://www.rstudio.com/products/rstudio/download/](https://www.rstudio.com/products/rstudio/download/){target="_blank"}

## Quarto and coding

3.  Now, open R Studio and create a new Quarto document with a title in the following format: "Firstname Lastname - HW0".

4.  Erase everything except the YAML header. Insert an R chunk like following, and run the following code in it. Make sure you customize it with your own name.

    ```{r}
    lastname = "Studio" #fill it
    firstname = "R" # fill it
    paste("My name is ", firstname, lastname)
    ```

5.  Also run the following code to see Air passengers data.

    ```{r}
    AirPassengers
    ```

6.  Run the following code, and comment on the display.

    ```{r}
    plot(AirPassengers)
    ```
